\documentclass[titlepage]{article}
\usepackage[left=2.54cm,top=2.54cm,right=2.54cm,nohead]{geometry}
\usepackage{graphicx}
\setlength{\parskip}{2mm}
\title{User Manual for OGSAS}
\author{Group 22 \\
\\
Yubin Kim (y12kim, 20239608) \\
Daniel Burstyn (dmbursty, 20206120) \\
Derek Thurn (dthurn, 20228984) \\
Nathaniel Flath (nflath, 20252804)
\\
\\for Divya Nair}
\begin{document}
\maketitle
\newpage
\tableofcontents
\newpage


\section{Introduction}
\subsection{Outline}
% Max
Max
\subsection{Product Overview}
% Yubin
Yubin
\subsection{First Sample Run}
% Nathan
Nathan


\section{Conventions}
text
\subsection{User Assumptions}
\begin{itemize}
\item User is a faculty member in the David R. Cheriton School of Computer
  Science at the University of Waterloo.
\item User is the Director of Admissions at the University of Waterloo.
\item User is logged in to the OGSAS system.
\item User is familiar with the University of Waterloo graduate admissions
  process.
\item User is familiar with basic computing paradigms, such as the mouse, the
  internet, and the keyboard.
\end{itemize}

\subsection{Notational Conventions}
text
\subsection{Terms}
% Max
\begin{itemize}
\item \textbf{Applicant} - A potential graduate student for the David R.
Cheriton School of Computer Science. They submit an application and are
allowed edit/view their supplementary information form and application check
list.
\item \textbf{Application Checklist} - A list of items to be completed or sent
in to UW for a complete application.
\item \textbf{Application Status} - The current state of an application. An
application initially starts as "New" and may be assigned to other statuses
such as "Hold", "Initial Review", "Circulate" at the discretion of the Admin
Staff.
\item \textbf{Reference Form} - A form that speaks of the Applicant's fitness
to be a graduate student, filled out by a referee. It is a part of the
Applicant's application.
\item \textbf{Supplementary Form} - A form composed of additional information
to be filled out by the Applicant.  It includes a statement about research
interests, resume, and academic background information.
\end{itemize}
\subsection{Acronyms and Abbreviations}
% Max
\begin{itemize}
\item GSO - \textbf{G}raduate \textbf{S}tudies \textbf{O}ffice
\item OUAC - \textbf{O}ntario \textbf{U}niversities' \textbf{A}pplication
\textbf{C}entre
\item UW - \textbf{U}niversity of \textbf{W}aterloo
\item UWID - \textbf{U}niversity of \textbf{W}aterloo \textbf{ID} Number
\end{itemize}


\section{Basic Use Cases}
text
\subsection{Administrative Tasks}
text
\subsubsection{Searching Applicants}
% Natt
Natt
\subsubsection{Viewing Application}
% Max
Max
\subsubsection{Changing the Status of an Application}
% Yubin
Y12
\subsubsection{Viewing the History of an Application}
% Thor
Thor
\subsection{Faculty Tasks}
text
\subsubsection{Changing your Opinion of an Application}
% Natt
Natt
\subsubsection{Commenting on an Application}
% Max
You can comment on a particular application if you find it necessary.  These
will only be visible to you and other faculty members.  To leave a
comment, navigate to the "Comments" tab on the application page, and write
your remarks in the large text box.  Once you submit your comment, it will be
timestamped and visible to all faculty members.
\subsubsection{Accepting an Application}
% Yubin
Yubink


\section{Advanced Features}
text
\subsection{Administrative Tasks}
text
\subsubsection{Viewing Statistics}
% Thor
Thor
\subsubsection{Viewing Access Roles}
% Natt
Natt
\subsubsection{Listing Authorized Users}
% Max
Max
\subsubsection{Uploading Scans}
% Yubin
Yubs
\subsubsection{Correcting Scores}
% Thor
Thor
\subsubsection{Changing an Applicant's Research Area}
% Thor
Thor
\subsubsection{Changing an Applicant's Term of Admission}
% Natt
Natt
\subsection{Faculty Tasks}
\subsubsection{Editing your Personal Info}
% Max
Max
\subsubsection{Recommending an Applicant to Another Faculty Member}
% Yubin
Kim


\section{Limitations}
\begin{itemize}
\item The system does not support direct communication between Faculty and the
  Applicant.
\item The system does not support the Applicant accepting an offer.
\item The system does not support uploading some data - such as TOEFL scores
  and transcripts - in computerized form.  Instead, these items must be
  scanned in.
  
\end{itemize}
\end{document}

% Local Variables: 
% fill-column:78
% End:
